### 2021-10-14 documentos/vector_space.tex
I'll explain when you're older!

Other two simple but crucial
observation are the following. For valid quantum channels, 
% \begin{equation}
% % \sum_{
% \end{equation}
$\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
$\lambda_\valpha=0$ for all $\valpha\in \Omega$, as each member of the sum 
greater or equal to zero, 
% $p_{\vec\alpha}+p_{\vec\beta}=0$ implies $p_{\vec\alpha}=p_{\vec\beta}=0$, as
% all $p_{\vec\alpha} \ge 0$
due to complete positivity of the underlying channel. 
Finally, looking into the first component of \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha):=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
\eref{eq:sum:alpha:is:2N} from 
% $\sum_{\vec\beta}p_{\vec\beta}=1$ (which follows from \cpnote{ver constante de 
% normalizacion chido y ver el motivo}), and 
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$ 
(which follows from the aforementioned assumption), 
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
% holds for all $\vec\beta\in\Phi(\vec\alpha)$. 
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three multi-indices, $\vec\alpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$. Let us now work out the nature
of this connection.

Since all elements of $A$ are $\pm1$, and using that $A_{\valpha \vbeta}=1$
of $\vbeta \in \Phi(\vec\alpha)$, we may rewrite \eref{eq:coneccion:dos:indices:A}
as
\begin{equation}
\san_{\vbeta\vec\gamma} \san_{\vbeta\vec{\gamma^\prime}} 
= 
\san_{\valpha\vbeta} 
, \quad
\forall \vec\beta\in\Phi(\vec\alpha),
\label{eq:12}
\end{equation}
which can be further decomposed in 
\begin{equation}
\sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}
  \ldots
%   \cdot\ldots\cdot 
\sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=
\sa_{\alpha_1\beta_1}
  \ldots
% \cdot\ldots\cdot
 \sa_{\alpha_N\beta_N}.
\label{eq:13}
\end{equation}
% Moreover, since 
% 
% 
% 
% Since $\sa_{\alpha\beta}=\pm1$, we may rewrite (\ref{eq:11}) as
% \begin{equation}
% \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=1.
% \label{eq:12}
% \end{equation}
% This must hold for all $\vec\beta\in\Phi(\vec\alpha)$. From this follows that
% (\ref{eq:12}) may be rewritten as
% \begin{equation}
% \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=
% \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}.
% \label{eq:13}
% \end{equation}
Since this must hold for all $\vec\beta\in\Phi(\vec\alpha)$, it follows that 
the above relation must hold individually for every index\cpnote{francois, recuerdo 
que esto lo discutimos. No se si en el momento me convencí, pero ahora no veo 
proque se tiene que cumplir para los indices individulaes}. Thus we have, 
that, for all $1\leq k\leq N$, $\gamma_k$
and $\gamma_k^\prime$ are connected by
\begin{equation}
\sa_{\beta\gamma_k}\sa_{\beta\gamma_k^\prime}=\sa_{\beta\alpha}
\label{eq:14}
\end{equation}
for all $0\leq\beta\leq3$. One can verify, by working out the different cases, 
that \eref{eq:14} equivalently expressed as 
\begin{equation}
\gamma_k^\prime=\alpha\oplus{}\gamma_k
\label{eq:oplus}
\end{equation}
where the $\oplus{}$ denotes the operation of the Kein group; see Figure
\ref{tab:1} for a detailed description. 
% \cpnote{Quiza poner una relacion como $\tau_{\vgamma}=\tau_{\valpha\oplus{}\vgamma}$}
\par
% \cpnote{by direct calculation I can verify 
% that its true, but I see it as a coincidence. Is they ere a say to know
% that \eref{eq:14} will lead to a sum?}
% }}}
% Espacio vectorial {{{

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$\cite{alguna:referencia}.  Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further \cpnote{desmechar} the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
\cpnote{maybe its not necesary}
}. 
We can indeed restate \eref{eq:oplus}, an say that for valid quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
% \cpnote{Creo que aca podemos referenciar la figura 2c y mostrar como
% funcionaria que no es cerrada bajo la suma}
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preseverd and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

@@

From this readily follows an amusing property: the set of all multi-indices
$\vec\gamma$ for which $\tau_{\vec\gamma}=1$, is closed under binary
vector addition, in other words, it forms a vector subspace of the set
of all multi-indices.
A moment's consideration will further show that the above reasoning can be
inverted, that is, that if we set all $\tau_{\vec\gamma}$ equal to one
whenever the $\vec\gamma$ belong to a given vector subspace of the set of
all indices, then the vector $\vec\tau$ indeed has an image which is a vector
with only positive components. 
In other words, there is a one to one correspondence between 
a valid quantum channel, and a vector subspace of the aforementioned space. 
\cpnote{Aca quizá que Jose Alfredo ponga un ejemplo de una 
figura en la que se tenga algo que es un canal y algo que no es  un canal. 
Quizá también se puede mencionar qeu el subespacio mas chico es el canal totalmente 
depolarizante y el subespacio mas grande es la identidad} \par

@@

The dimension of the vector space $V$ of all multi-indices is $2N$, and we have
seen earlier that
\cpnote{Why is the dimension $2N$? I can work it out, but just after adding one 
by one elements}
\begin{equation}
W=\left\{
\vec\alpha:\vec\alpha\in V, \tau_{\vec\alpha}=1
\right\}
\label{eq:19}
\end{equation}
is a subspace of $V$. As such, $W$ has a given dimension $K$, which means that
$W$ has $2^K$ elements. In other words, a vector $\vec\tau$ with the property
discussed above can only have $2^K$ elements equal to 1, for a given integer
$K$.

\cpnote{Dejar que la raza elija uno de los dos parrafos anteriores}
 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCEs exist that conserve $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest an relation  between individual channels that for the time being
has escaped our efforts to identify. 
\par % }}}
% Frase de David {{{
\ddnote{Agrego esto:} A final remarks of the derivation as a tool. Concepts
might be complicated, but the tool gives some simple rules to the reader:
\textit{tell me what components you want to keep, and the (vector) summation
rules will give you the rest of components to make the target operation a CPTP
one.}
\par % }}}

>>>>

Other two simple but crucial
observation are the following. For valid quantum channels, 
% \begin{equation}
% % \sum_{
% \end{equation}
$\sum_{\valpha \in \Omega} \lambda_\valpha=0$ implies that 
$\lambda_\valpha=0$ for all $\valpha\in \Omega$, as each member of the sum 
greater or equal to zero, 
% $p_{\vec\alpha}+p_{\vec\beta}=0$ implies $p_{\vec\alpha}=p_{\vec\beta}=0$, as
% all $p_{\vec\alpha} \ge 0$
due to complete positivity of the underlying channel. 
Finally, looking into the first component of \eref{eq:tau:is:A:p}, and taking into 
account the normalization condition that $\tau_{\vec 0}=1$, we obtain
\begin{equation}
\sum_\valpha \lambda_\valpha = 2^N,
\label{eq:sum:alpha:is:2N}
\end{equation}
since, $A_{\vec 0, \vbeta}=1$ for all $\vbeta$. \par
% }}}
% Conexion de los indices para llegar al grupo de Klein {{{
Now we need a definition: to each multi-index $\vec\alpha$ we associate a \textit{set}
of multi-indices $\Phi(\vec\alpha)$ as follows
\begin{equation}
\Phi(\vec\alpha):=\left\{
\vec\beta:\san_{\vec\alpha\vec\beta}=1
\right\}
% =\left\{ \vec\beta: \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}
% =1
% \right\}
\label{eq:8}
\end{equation}
If we now assume that $\tau_{\vec\alpha}=1$, it follows from subtracting 
\eref{eq:sum:alpha:is:2N} from 
$\sum_{\vec\beta}\san_{\vec\alpha\vec\beta}\lambda_{\vec\beta}=2^N$ 
(which follows from the aforementioned assumption), 
% \begin{subequations}
% \begin{eqnarray}
% \sum_{\vec\beta}p_{\vec\beta}&=&1
% \label{eq:9a}\\
% \sum_{\vec\beta}\san_{\vec\alpha\vec\beta}p_{\vec\beta}&=&\tau_{\vec\alpha}=1
% \end{eqnarray}
% \label{eq:9}
% \end{subequations}
% that for all $\vec\beta\notin\Phi(\vec\alpha)$
\begin{eqnarray}
\lambda_{\vec\beta}=0,\quad \forall \vec\beta\notin\Phi(\vec\alpha).
\label{eq:condition:lambda:zero:Phi}
\end{eqnarray}
Thus, if $\tau_{\vec\alpha}=1$, then $\tau_{\vec\gamma}$
and $\tau_{\vec{\gamma^\prime}}$ are equal if
% , for all $\vec\beta\in\Phi(\vec\alpha)$.
\begin{equation}
\san_{\vbeta\vec\gamma}=\san_{\vbeta\vec{\gamma^\prime}}, \quad
\forall \vec\beta\in\Phi(\vec\alpha).
\label{eq:coneccion:dos:indices:A}
\end{equation}
% holds for all $\vec\beta\in\Phi(\vec\alpha)$. 
This follows from restricting the sum \eref{eq:tau:is:A:p}
to the $\lambda_\vbeta \ne 0$, given in \eref{eq:condition:lambda:zero:Phi}. 
Condition (\ref{eq:coneccion:dos:indices:A}) therefore connects three multi-indices, $\vec\alpha$,
$\vec\gamma$ and $\vec{\gamma^\prime}$.
When such a connection exists, $\tau_{\vec\alpha}=1$ implies
$\tau_{\vec\gamma}=\tau_{\vec{\gamma^\prime}}$. Let us now work out the nature
of this connection.

Since all elements of $A$ are $\pm1$, and using that $A_{\valpha \vbeta}=1$
of $\vbeta \in \Phi(\vec\alpha)$, we may rewrite \eref{eq:coneccion:dos:indices:A}
as
\begin{equation}
\san_{\vbeta\vec\gamma} \san_{\vbeta\vec{\gamma^\prime}} 
= 
\san_{\valpha\vbeta} 
, \quad
\forall \vec\beta\in\Phi(\vec\alpha),
\label{eq:12}
\end{equation}
which can be further decomposed in 
\begin{equation}
\sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}
  \ldots
%   \cdot\ldots\cdot 
\sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=
\sa_{\alpha_1\beta_1}
  \ldots
% \cdot\ldots\cdot
 \sa_{\alpha_N\beta_N}.
\label{eq:13}
\end{equation}
% Moreover, since 
% 
% 
% 
% Since $\sa_{\alpha\beta}=\pm1$, we may rewrite (\ref{eq:11}) as
% \begin{equation}
% \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=1.
% \label{eq:12}
% \end{equation}
% This must hold for all $\vec\beta\in\Phi(\vec\alpha)$. From this follows that
% (\ref{eq:12}) may be rewritten as
% \begin{equation}
% \sa_{\beta_1\gamma_1}\sa_{\beta_1\gamma_1^\prime}\cdot\ldots\cdot \sa_{\beta_N\gamma_N}\sa_{\beta_N\gamma_N^\prime}=
% \sa_{\alpha_1\beta_1}\cdot\ldots\cdot \sa_{\alpha_N\beta_N}.
% \label{eq:13}
% \end{equation}
Since this must hold for all $\vec\beta\in\Phi(\vec\alpha)$, it follows that 
the above relation must hold individually for every index\cpnote{francois, recuerdo 
que esto lo discutimos. No se si en el momento me convencí, pero ahora no veo 
proque se tiene que cumplir para los indices individulaes}. Thus we have, 
that, for all $1\leq k\leq N$, $\gamma_k$
and $\gamma_k^\prime$ are connected by
\begin{equation}
\sa_{\beta\gamma_k}\sa_{\beta\gamma_k^\prime}=\sa_{\beta\alpha}
\label{eq:14}
\end{equation}
for all $0\leq\beta\leq3$. One can verify, by working out the different cases, 
that \eref{eq:14} equivalently expressed as 
\begin{equation}
\gamma_k^\prime=\alpha\oplus{}\gamma_k
\label{eq:oplus}
\end{equation}
where the $\oplus{}$ denotes the operation of the Kein group; see Figure
\ref{tab:1} for a detailed description. 
\par
% }}}
% Espacio vectorial {{{

It will be useful to think of the multi-index $\valpha$ as an element of 
a vector space. To do so, we notice that any group with the property that
$\alpha \oplus \alpha =0$ is indeed a vector space under the two-element field
$\{0,1\}$\cite{alguna:referencia}.  Then, we build the complete vector space,
with the same field, and defining $\valpha \oplus \vbeta = (\alpha_1 \oplus
\beta_1,\cdots, \alpha_N \oplus \beta_N)$
\footnote{
We can further \cpnote{desmechar} the vector space, by identifying each index
$\alpha \in \{0,1,2,3\}$ with its binary notation [e.g. identify $2$ with $(1,0)$]
such that to each multindex  $\valpha$ there corresponds a binary string of 
length $2N$. In fact, our sum $\oplus$ would correspond to addition modulo 2 of 
each of the $2N$ components. 
}. 
We can indeed restate \eref{eq:oplus}, an say that for valid quantum channels, 
if $\tau_{\vec\alpha}=1$, then
$\tau_{\vec\gamma}=\tau_{\vec\alpha\oplus{}\vec\gamma}$.
For example, in \fref{fig:two:qubit:examples}(c) the indices that 
correspond to preserved components are 
$\valpha^{(0)} = (0,0)$, 
$\valpha^{(1)} = (0,2)$, 
$\valpha^{(2)} = (1,0)$, 
$\valpha^{(3)} = (2,2)$, and
$\valpha^{(4)} = (3,2)$. However, 
$\valpha^{(1)} + \valpha^{(2)} = (1,2)$, which is not preseverd and thus this 
diagram does not correspond to a quantum channel.  From this view 
we can derive several interesting observations that will be presented in 
the rest of the section.

@@

From this readily follows an amusing property: the set of all multi-indices
$\vec\gamma$ for which $\tau_{\vec\gamma}=1$, is closed under binary
vector addition, in other words, it forms a vector subspace of the set
of all multi-indices.
A moment's consideration will further show that the above reasoning can be
inverted, that is, that if we set all $\tau_{\vec\gamma}$ equal to one
whenever the $\vec\gamma$ belong to a given vector subspace of the set of
all indices, then the vector $\vec\tau$ indeed has an image which is a vector
with only positive components. 
In other words, there is a one to one correspondence between 
a valid quantum channel, and a vector subspace of the aforementioned space. 
% \cpnote{Aca quizá que Jose Alfredo ponga un ejemplo de una 
% figura en la que se tenga algo que es un canal y algo que no es  un canal. 
% Quizá también se puede mencionar qeu el subespacio mas chico es el canal totalmente 
% depolarizante y el subespacio mas grande es la identidad} 
\par

@@

The dimension of the vector space $V$ of all multi-indices is $2N$, and we have
seen earlier that
\begin{equation}
W=\left\{
\vec\alpha:\vec\alpha\in V, \tau_{\vec\alpha}=1
\right\}
\label{eq:19}
\end{equation}
is a subspace of $V$. As such, $W$ has a given dimension $K$, which means that
$W$ has $2^K$ elements. In other words, a vector $\vec\tau$ with the property
discussed above can only have $2^K$ elements equal to 1, for a given integer
$K$. \par

\cpnote{Dejar que la raza elija uno de los dos parrafos anteriores} \par

 % }}}
% Cuantos canales conservan $2^K$ componentes? {{{
It is natural to ask how many PCEs exist that conserve $2^K$ components. One
can calculate such number, $\sss_{N,K}$, by examining the number of different
independent subsets of vectors that spawn a given vector subspace. In appendix 
\ref{app:contar} we show that 
\begin{equation}
\sss_{N,K}=\prod_{m=0}^{K-1} \frac{2^{2N-m}-1}{2^{K-m}-1}.
\label{eq:conteo:main}
\end{equation}
%An elementary test is $N=3$ and $K=2,3$: 
%\begin{eqnarray}
%\sss_{3,2}&=&\frac{(2^6-1)(2^5-1)}{(2^2-1)(2^1-1)}=\frac{63\cdot31}{3}=651,\\
%\sss_{3,3}&=&\frac{(2^6-1)(2^5-1)(2^4-1)}{(2^3-1)(2^2-1)(2^1-1)}=\frac{63\cdot31\cdot15}{7\cdot3}=1395,
%\end{eqnarray}
%which are indeed the values found numerically. 
From the above expression, it is easy to see the symmetry relation
\begin{equation}
\sss_{N,K}=\sss_{N,2N-K}
\end{equation}
which suggest an relation  between individual channels that for the time being
has escaped our efforts to identify. 
\par % }}}
% Frase de David {{{
\ddnote{Agrego esto:} A final remarks of the derivation as a tool. Concepts
might be complicated, but the tool gives some simple rules to the reader:
\textit{tell me what components you want to keep, and the (vector) summation
rules will give you the rest of components to make the target operation a CPTP
one.}
\par % }}}
